Language-Based Low-Level MPC Security: $\metaprot$ to $\minifed$

### 2024-05-09 ppdp24/bruteforce.tex
checkpoint

We define the syntax of Datalog as follows. As per standard
nomenclature, \emph{atoms} are $\minifed$ variables $x$,
\emph{literals} are atoms or negated atoms, and clause bodies are
conjunctions of literals.  A \emph{fact} is a clause with no body. A
\emph{Datalog program} is a list of clauses.
$$
\begin{array}{rclr}
  \mathit{body} &::=&  x \mid \neg x \mid x \wedge \mathit{body} \mid \neg x \wedge \mathit{body} \mid \varnothing \\
  \mathit{clause} &::=& x \gets \mathit{body} \mid x \gets
\end{array}
$$
When translating protocols, we need to extract the variables
that occur in expressions $\be$ computed by a client $\cid$,
written $\vars\ \be\ \cid$ where:
\begin{mathpar}
  \vars\ \secret{w} \cid \defeq \{ \elab{\secret{w}}{\cid} \}

@@

In addition to converting view definitions to logic clauses, we also need to convert
secrets and random tapes. Since we assume given values for these in an arbitrary run of
the program, we can capture these as a particular fact base.
\begin{definition}
  Given $\store$, let $\{x_1,\ldots,x_n \} =
  \{ x \in \dom(\store) \mid \store(x) = 1 \}$.
  Then define $\mathit{facts}(\store) \defeq x_1 \gets, \ldots, x_n \gets$.
\end{definition}
The following result ties these pieces together and establishes
correctness of this approach.
\begin{lemma}
  For all $\prog$, 
  $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
  program \cite{aspis2018linear}, and $\store$ is the unique Least Herbrand
  Model of: $$(\mathit{facts}(\store_{\secrets(\prog) \cup \flips(\prog)}),\datalog(\prog))$$
  iff $\store \in \runs(\prog)$.
\end{lemma}
Finally, to compute $\runs(\prog)$, and thus $\progtt(\prog)$, we compute
the Least Herbrand Model $\store$ of $(\mathit{facts}(\store'),\datalog(\prog))$
for all $\store' \in \mems(\secrets(\prog) \cup \flips(\prog))$, observing
that model computation for distinct fact bases can be done in parallel. 
We have used this method to verify the one-time-pad encryption and Shamir
addition protocols in Section \ref{section-lang-example}, as well as
larger protocol components as we discuss in Section \ref{section-verification}.

>>>>

\begin{comment}
We immediately note that $\runs(\prog)$ can be obtained by a left-folding
of $\solvealg$ across $\prog$. 
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\xassign{x}{\be}{\cid}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}
\end{comment}

We define the syntax of Datalog as follows. As per standard
nomenclature, \emph{atoms} are $\minifed$ variables $x$,
\emph{literals} are atoms or negated atoms, and clause bodies are
conjunctions of literals.  A \emph{fact} is a clause with no body. A
\emph{Datalog program} is a list of clauses.
$$
\begin{array}{rclr}
  \mathit{body} &::=&  x \mid \neg x \mid x \wedge \mathit{body} \mid \neg x \wedge \mathit{body} \\
  \mathit{clause} &::=& x \gets \mathit{body} \mid x \gets
\end{array}
$$
When translating protocols, we need to extract the variables
that occur in expressions $\be$ computed by a client $\cid$,
written $\vars\ \be\ \cid$ where:
\begin{mathpar}
  \vars\ \secret{w} \cid \defeq \{ \elab{\secret{w}}{\cid} \}

@@

In addition to converting view definitions to logic clauses, we also need to convert
secrets and random tapes. Since we assume given values for these in an arbitrary run of
the program, we can capture these as a particular fact base.
\begin{definition}
  Given $\store$, let $\{x_1,\ldots,x_n \} =
  \{ x \in \dom(\store) \mid \store(x) = 1 \}$.
  Then define $\mathit{facts}(\store) \defeq x_1 \gets, \ldots, x_n \gets$.
\end{definition}
The following result ties these pieces together and establishes
correctness of this approach.
\begin{lemma}
  For all $\prog$, 
  $\datalog(\prog)$ is a \emph{normal}, \emph{stratified}
  program \cite{aspis2018linear}, and $\store$ is the unique Least Herbrand
  Model of: $$(\mathit{facts}(\store_{\secrets(\prog) \cup \flips(\prog)}),\datalog(\prog))$$
  iff $\store \in \runs(\prog)$.
\end{lemma}
Finally, to compute $\runs(\prog)$, and thus $\progtt(\prog)$, we compute
the Least Herbrand Model $\store$ of $(\mathit{facts}(\store'),\datalog(\prog))$
for all $\store' \in \mems(\secrets(\prog) \cup \flips(\prog))$, observing
that model computation for distinct fact bases can be done in parallel.

We can then query representations of $\runs(\prog)$ with
implementations of conditioning and marginalization to automatically
verify the passive model hyperproperties described in Section
\ref{section-hyper}-- in particular correctness and noninterference
modulo output, which imply passive security.  We have used this method
to verify security in $\mathbb{F}_2$ of Shamir addition as defined in Section
\ref{section-lang-example}, and single-gate $\eand$ and $\exor$ circuits with GMW
Beaver triples. We have also
verified properties compositional properties of gates themselves that establish
circuit invariants that imply passive security in larger circuits.
We discuss these examples and proof methods in Section \ref{section-examples}.

### 2024-05-09 ppdp24/examples.tex
checkpoint

\section{Examples}

\subsection{Example: 2-Party GMW}
\label{section-metalang-gmw}

@@

>>>>

\section{Examples}
\label{section-examples}

\subsection{2-Party GMW}
\label{section-metalang-gmw}

@@

\subsection{2-Party BDOZ}

\input{bdoz}

\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$,
  let $X_H \subseteq \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$.
  Then the \emph{adversarial inputs to $X_H$} is the least set
  $X_C \subseteq \cinputs$ such that $\progtt(\prog) \not\vdash X_C * X_H$.
\end{definition}
\begin{definition}[Cheating Detection]
  \emph{Cheating is detected} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
  for all  $\store \in \aruns(\prog)$,
  letting $X_H = \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$,
  and letting $X_C$ be the adversarial inputs to $X_H$,
  there exists $\sigma'\in \runs(\prog)$
  with $\store_{X_C} = \store'_{X_C}$.  
\end{definition}

\begin{lemma}
  If cheating is detected in $\prog$, then $\prog$ has integrity.
\end{lemma}

### 2024-05-09 ppdp24/gmw.tex
checkpoint

\begin{fpfig}[t]{2-Party GMW circuit library with And gate (T), and type annotations (B).}{fig-gmw}
{\footnotesize
  \begin{verbatimtab}
    encodegmw(in1,in2) {
      v[1, in1++"out"] := s[1,in1] xor flip[1,in1];
      v[2, in1++"out"] := flip[1,in1];
      v[1, in2++"out"] := s[2,in2] xor flip[2,in2];
      v[2, in2++"out"] := flip[2,in2]
      { shares1 = { c1 = v[1, in1++"out"]; c2 = v[2, in1++"out"] };
        shares2 = { c1 = v[1, in2++"out"]; c2 = v[2, in2++"out"]} } 
    }

>>>>

\begin{fpfig}[t]{2-Party GMW circuit library with And gate.}{fig-gmw}
{\footnotesize
  \begin{verbatimtab}
    encodegmw(in1,in2) {
      m[in1++"out"]@1 := (s[in1] xor r[in1])@1;
      m[in1++"out"]@2 := r[in1]@1;
      m[in2++"out"]@1 := (s[in2] xor flip[in2])@2;
      m[in2++"out"]@2 := r[in2]@2
      { shares1 = { c1 = v[1, in1++"out"]; c2 = v[2, in1++"out"] };
        shares2 = { c1 = v[1, in2++"out"]; c2 = v[2, in2++"out"]} } 
    }

### 2024-05-09 ppdp24/metalang.tex
checkpoint

The $\fedprot$ language is low-level and does not include abstractions for
defining composable elements. So in this Section we introduce the $\metaprot$
language which that includes structured data and function definitions
for defining composable protocol elements at a higher level of abstraction. 
The $\metaprot$ language is a \emph{metalanguage}, where 
$\fedprot$ protocols are the residuum of computation.

In addition to these declarative benefits of $\metaprot$, component
definitions support compositional verification of larger programs.
The techniques in Section \ref{section-bruteforce} are not feasible
for large boolean circuits, for example, but they are for verification
of properties of compositional units, as we discuss in Section
\ref{section-verification}.

@@

The syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot}.  It includes a syntax of function
definitions and records, and values $\mv$ include client ids $\cid$, identifier
strings $w$, expressions $\be$ in field $\mathbb{F}_p$, and the unit value $()$.
Expression forms allow dynamic construction of field expressions and $\minifed$ commands.
The construction of a command has the side-effect of adding to the residual
$\minifed$ protocol. Formally, we consider a complete metaprogram to include both a
codebase and a ``main'' program that uses the codebase. 
\begin{definition}
A \emph{codebase} $\codebase$ is a list of function 
declarations. We write $ \codebase(f) = y_1,\ldots,y_n,\ e$
iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
codebase and expression $\codebase, e$. We may omit
$\codebase$ if it is clear from context.  
\end{definition}

When we consider larger examples, our
focus will be on developing a codebase that can be used to define
arbitrary circuits, i.e., complete and concrete protocols. Since
strings and identifiers can be constructed manually, and expressions
can occur inside assignments and field expression forms, function
definitions can generalize over $\fedprot$-level patterns to obtain
composable program units.

\subsection{Semantics}

>>>>

The $\fedprot$ language is low-level and does not include abstractions for
defining composable elements. So in this Section we introduce the $\metaprot$
language which that includes structured data and function definitions
for defining composable protocol elements at a higher level of abstraction. 
The $\metaprot$ language is a \emph{metalanguage}, where 
$\fedprot$ protocols are the residuum of computation. In addition to these
declarative benefits of $\metaprot$, component
definitions support compositional verification of larger programs as we will
discuss with examples in Section
\ref{section-examples}.

@@

The syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot}.  It includes a syntax of function
definitions and records, and values $\mv$ include client ids $\cid$, identifier
strings $w$, expressions $\be$ in field $\mathbb{F}_p$, and the unit value $()$.
Expression forms allow dynamic construction of field expressions and $\minifed$ commands.
The construction of a command has the side-effect of adding to the residual
$\minifed$ protocol. Formally, we consider a complete metaprogram to include both a
codebase and a ``main'' program that uses the codebase. 
\begin{definition}
A \emph{codebase} $\codebase$ is a list of function 
declarations. We write $ \codebase(f) = y_1,\ldots,y_n,\ e$
iff $f(y_1,\ldots,y_n) \{ e \} \in \codebase$.
A \emph{metaprogram}, aka \emph{metaprotocol}  is a pair of a 
codebase and expression $\codebase, e$. We may omit
$\codebase$ if it is clear from context.  
\end{definition}

When we consider larger examples, our
focus will be on developing a codebase that can be used to define
arbitrary circuits, i.e., complete and concrete protocols. Since
strings and identifiers can be constructed manually, and expressions
can occur inside assignments and field expression forms, function
definitions can generalize over $\fedprot$-level patterns to obtain
composable program units.
\subsection{Semantics}
